---
title: "01_Data-reorganisation"
format: html
editor: visual
---

## Load libraries 

```{r load-libraries}
#1 message : false
library(tidyverse)
library(ez)
```

## Load the data

```{r load-data}

# Read a data file that is text with columns separated by spaces and strings as factors

Results <- read.table("../data/Results.txt", header = TRUE, sep = " ", stringsAsFactors = TRUE)

```

# 1. Data reorganization

Real data are often messy. Before running the ANOVA, we check that the design is complete (the same number of measures in every condition) and we remove the conditions that are incomplete. We use the tidyverse library for this.

### Count the measures by GROUP and ID

We count how many data points we have for each combination of GROUP and ID. If a combination has fewer measures than the others, some data are missing.

```{r Count-measures-by-group-id}
Results %>%
  count(GROUP, ID)
```

**Observation:** data are missing for ID3 in the group pp (400 measures instead of 760). We therefore remove ID3 from the analysis. 

### Remove ID3 and recount

filter(ID != "ID3") keeps all the rows except those where ID is ID3.

```{r Remove-ID3-and-recount}
Results %>%
  filter(ID != "ID3") %>%
  count(GROUP, ID)
```

The counts still differ between groups, but this is expected: the groups do not have the same number of participants. We still have a problem, so we check the other factors one after the other.

### Check the number of measures by REP

```{r Count-measures-by-rep}

Results %>%
  filter(ID != "ID3") %>%
  count(REP)

```

**Observation:** R3 has fewer measures than R1 and R2. We look at which groups are responsible.

### Remove REP3 and recount

```{r Remove-REP3-and-recount}

Results %>%
  filter(ID != "ID3") %>%
  filter(REP != "R3") %>%
  count(GROUP, REP)

```

Now every group has the same number of measures in R1 and R2: the design is clean.

















