---
title: "01_Data-reorganisation"
format: html
editor: visual
---

## Load libraries 

```{r load-libraries}
#1 message : false
library(tidyverse)
library(ez)
```

## Load the data

```{r load-data}

# Read a data file that is text with columns separated by spaces and strings as factors

Results <- read.table("../data/Results.txt", header = TRUE, sep = " ", stringsAsFactors = TRUE)

```

# 1. Data reorganization

Real data are often messy. Before running the ANOVA, we check that the design is complete (the same number of measures in every condition) and we remove the conditions that are incomplete. We use the tidyverse library for this.

### Count the measures by GROUP and ID

We count how many data points we have for each combination of GROUP and ID. If a combination has fewer measures than the others, some data are missing.

```{r Count-measures-by-group-id}
Results %>%
  count(GROUP, ID)
```

**Observation:** data are missing for ID3 in the group pp (400 measures instead of 760). We therefore remove ID3 from the analysis. 

### Remove ID3 and recount

filter(ID != "ID3") keeps all the rows except those where ID is ID3.

```{r Remove-ID3-and-recount}
Results %>%
  filter(ID != "ID3") %>%
  count(GROUP, ID)
```

The counts still differ between groups, but this is expected: the groups do not have the same number of participants. We still have a problem, so we check the other factors one after the other.

### Check the number of measures by REP

```{r Count-measures-by-rep}

Results %>%
  filter(ID != "ID3") %>%
  count(REP)

```

**Observation:** R3 has fewer measures than R1 and R2. We look at which groups are responsible.

### Remove REP3 and recount

```{r Remove-REP3-and-recount}

Results %>%
  filter(ID != "ID3") %>%
  filter(REP != "R3") %>%
  count(GROUP, REP)

```

Now every group has the same number of measures in R1 and R2: the design is clean.

### Count the participants per group

n_distinct(SUBJ) counts the different participants, not the number of rows.

```{r Count-the-number-of-participant-per-group}

Results %>%
  filter(ID != "ID3") %>%
  filter(REP != "R3") %>%
  group_by(GROUP) %>%
  summarise(count = n_distinct(SUBJ))

```

We have 13 participants in ap, 19 in cp and 19 in pp (51 in total).

### Create the clean dataframe

We save the cleaned data in a new object clean_results, so we do not have to repeat the filters every time. droplevels() removes the levels ID3 and R3 that the factors still remember after filtering.

```{r Create-clean-dataframe}

clean_results <- Results %>%
  filter(ID != "ID3") %>%
  filter(REP != "R3") %>%
  droplevels()

```

## Run the ANOVA

We want to know whether the effect of task difficulty (ID) on movement time differs between groups, i.e., whether there is a **GROUP × ID interaction**.

The design is mixed:

- GROUP is a between-subjects factor (different participants in each group).
- ID is a within-subjects factor (each participant performs both difficulty levels).

We use ezANOVA() from the ez library. wid identifies the participant, and type = 3 is the type of sum of squares (it matters because the groups are unbalanced).

```{r Run-the-ANOVA}

res_anova <- ezANOVA(
  data     = clean_results,
  dv       = MovementTime,
  wid      = SUBJ,
  between  = GROUP,
  within   = ID,
  type     = 3,
  detailed = FALSE
)

res_anova

```

## Plot the interaction between GROUP and ID

To visualize the interaction, we plot the mean movement time of each GROUP as a function of ID. The lines connect the means, the points show the means, and the error bars show the mean ± 0.5 standard deviation.

### Interaction plot reference

```{r Interaction-plot-reference}


interaction_plot_reference <- ggplot(clean_results) +
  aes(x = ID, y = MovementTime, color = GROUP) +
  # lines: one line type per group
  stat_summary(fun = mean, geom = "line",
               aes(group = GROUP, linetype = GROUP), alpha = 0.7) +
  # error bars: mean ± 0.215 s, same size for all groups
  stat_summary(fun.data = function(x) data.frame(y    = mean(x),
                                                 ymin = mean(x) - 0.215,
                                                 ymax = mean(x) + 0.215),
               geom = "errorbar", width = 0.25, alpha = 0.7) +
  # points: one shape per group
  stat_summary(fun = mean, geom = "point",
               aes(shape = GROUP), size = 1.5, alpha = 0.7) +
  labs(x = "ID", y = "Movement Time (s)")

show(interaction_plot_reference)
```

### Interaction plot with within (group x ID) standard deviation

If the lines were parallel, there would be no interaction. Lines with different slopes indicate that the effect of ID depends on the group.

```{r Interaction-plot}
# plot the interaction between GROUP and ID and add error bars containing the 
# within (group x ID) standard deviation around the mean
interaction_plot_Group_ID <- ggplot(clean_results) +
  aes(x = ID, y = MovementTime, color = GROUP) +
  stat_summary(fun = mean, geom = "line", aes(group = GROUP)) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = .5), geom = "errorbar", width = 0.01) +
  stat_summary(fun = mean, geom = "point", shape= 21, fill = "white", size = 3) +
  labs(title = "Interaction between GROUP and ID on MovementTime",
       x = "ID",
       y = "Mean MovementTime (s)") +
  theme_minimal()

show(interaction_plot_Group_ID)
```

## Conclusion

The GROUP × ID interaction is significant, F(2, 48) = 11.58, p < .001. The effect of task difficulty on movement time differs between groups: movement time increases much more with ID in group pp than in groups ap and cp.
















