## Linear regression by GROUP

The ANOVA treats `ID1` and `ID2` as two labels. The article goes one step further and describes each group by a line, $MT = a + b \times ID$ (Fitts's law), where ID is a number of bits, the slope $b$ is the time cost of one more bit of difficulty and the intercept $a$ is the predicted time for a movement without accuracy constraint [@mottet2017]. We compute this line for each group.

### ID as a number

A regression needs a numerical predictor. In the article, the two levels of difficulty are 3.80 and 5.43 bits [@mottet2017, Methods]. The conversion is done once, in `R/setup.R` (column `ID_num`, shown in *Setup*). We check that each label received the right value:

```{r}
#| label: check-id-num
clean_results %>%
  count(ID, ID_num)
```

### One regression per group

The function below fits the line for each group of a data frame and returns the intercept, the slope and the equation as text. We wrote a function because we fit the same model on other versions of the data in section 1.7.

```{r}
#| label: fit-by-group
fit_by_group <- function(data) {
  data %>%
    group_by(GROUP) %>%
    group_modify(~ {
      model <- lm(MovementTime ~ ID_num, data = .x)
      tibble(intercept = coef(model)[[1]], slope = coef(model)[[2]])
    }) %>%
    ungroup() %>%
    mutate(equation = sprintf("MT = %.3f + %.3f * ID", intercept, slope))
}

regression_by_group <- fit_by_group(clean_results)

knitr::kable(regression_by_group, digits = 3)
```

These are the three equations given in the series, so our cleaning and our conversion of ID reproduce the expected result.

### What these equations do and do not tell us

```{r}
#| label: slopes
#| include: false
slope <- with(regression_by_group, setNames(slope, GROUP))
```

**What they tell us.** The slope is positive in the three groups: movement time increases with difficulty. One more bit costs `r round(slope["pp"], 2)` s to the patients, against `r round(slope["ap"], 2)` s and `r round(slope["cp"], 2)` s to the age-matched and young controls, that is `r round(slope["pp"] / slope["ap"], 1)` and `r round(slope["pp"] / slope["cp"], 1)` times more. The difference between the slopes is the GROUP $\times$ ID interaction of section 1.3, expressed in seconds per bit.

**What they do not tell us.** We see three limits.

- *The regression adds no information to the two means.* With only two values of ID, the least-squares line of a group passes exactly through its two mean movement times: the slope is (mean at ID2 $-$ mean at ID1) / (5.43 $-$ 3.80). We cannot test whether the relation is really linear, a limit that the article also states. The file does contain a third level, but not all patients did it and we do not know its value in bits.
- *The intercept is an extrapolation.* It is the predicted movement time at ID = 0, which is 3.8 bits away from the nearest measurement, so a small change in the two means is amplified. The negative intercept of the patients cannot be a movement time; it only says where the extended line crosses the vertical axis.
- *We report coefficients, not their uncertainty.* The model is fitted on all movements as if they were independent, although they come from `r n_distinct(clean_results$SUBJ)` participants. This does not change the intercepts and slopes, but the standard errors and p-values that `lm()` would print are too optimistic, so we do not report them. Our test of the difference between groups is the ANOVA, which is based on participants.
