## Linear regression by GROUP

The ANOVA treats `ID1` and `ID2` as two labels. The article goes one step further and describes each group by a line, $MT = a + b \times ID$ (Fitts's law), where ID is a number of bits, the slope $b$ is the time cost of one more bit of difficulty and the intercept $a$ is the predicted time for a movement without accuracy constraint [@mottet2017]. We compute this line for each group.

### ID as a number

A regression needs a numerical predictor. In the article, the two levels of difficulty are 3.80 and 5.43 bits [@mottet2017, Methods]. The conversion is done once, in `R/setup.R` (column `ID_num`, shown in *Setup*). We check that each label received the right value:

```{r}
#| label: check-id-num
clean_results %>%
  count(ID, ID_num)
```

### One regression per group

The function below fits the line for each group of a data frame and returns the intercept, the slope and the equation as text. We wrote a function because we fit the same model on other versions of the data in section 1.7.

```{r}
#| label: fit-by-group
fit_by_group <- function(data) {
  data %>%
    group_by(GROUP) %>%
    group_modify(~ {
      model <- lm(MovementTime ~ ID_num, data = .x)
      tibble(intercept = coef(model)[[1]], slope = coef(model)[[2]])
    }) %>%
    ungroup() %>%
    mutate(equation = sprintf("MT = %.3f + %.3f * ID", intercept, slope))
}

regression_by_group <- fit_by_group(clean_results)

knitr::kable(regression_by_group, digits = 3)
```

These are the three equations given in the series, so our cleaning and our conversion of ID reproduce the expected result.

### What these equations do and do not tell us

```{r}
#| label: slopes
#| include: false
slope <- with(regression_by_group, setNames(slope, GROUP))
```

**What they tell us.** The slope is positive in the three groups: movement time increases with difficulty. One more bit costs `r round(slope["pp"], 2)` s to the patients, against `r round(slope["ap"], 2)` s and `r round(slope["cp"], 2)` s to the age-matched and young controls, that is `r round(slope["pp"] / slope["ap"], 1)` and `r round(slope["pp"] / slope["cp"], 1)` times more. The difference between the slopes is the GROUP $\times$ ID interaction of section 1.3, expressed in seconds per bit.
