## Graph of movement time as a function of ID

The goal is a graph that resembles Fig. 3 (left panel) of the article: for each group, the mean movement time at the two levels of difficulty, error bars, the regression line and its equation.

Everything we need already exists: the means and standard deviations are in `cell_means` (section 1.3) and the lines and equations in `regression_by_group` (section 1.4). This section only draws.

### Choices we made

- **Error bars.** The caption of Fig. 3 says that the bars are standard deviations between participants. We use the `sd` column of `cell_means`.
- **Lines.** The regression lines are drawn over the whole x axis, as in the article.
- **Axes.** The x axis starts at 0 so that the intercepts are visible, and the y axis goes below 0 because the line of the patients does.
- **Equations.** Their text comes from `regression_by_group`, so the graph cannot disagree with the table of section 1.4.

```{r}
#| label: fig-mt-id
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Movement time as a function of the index of difficulty. Points: group means; bars: one standard deviation between participants; dashed lines: linear regression of each group."

# Position of the three equations, one under the other
equation_labels <- regression_by_group %>%
  mutate(x = 0.3, y = 6.9 - 0.4 * row_number())

graph_MT_ID <- ggplot(cell_means) +
  aes(x = ID_num, y = mean, color = GROUP) +
  geom_abline(
    data = regression_by_group,
    aes(intercept = intercept, slope = slope, color = GROUP),
    linetype = "dashed",
    show.legend = FALSE
  ) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    width = 0
  ) +
  geom_point(
    shape = 21,
    fill = "white",
    size = 3
  ) +
  geom_text(
    data = equation_labels,
    aes(x = x, y = y, label = equation),
    hjust = 0,
    show.legend = FALSE
  ) +
  scale_x_continuous(breaks = 0:6) +
  coord_cartesian(
    xlim = c(0, 6.2),
    ylim = c(-1, 7.5),
    expand = FALSE
  ) +
  labs(
    x = "ID (bit)",
    y = "Movement time (s)"
  )

graph_MT_ID
```

### Reading the graph

@fig-mt-id has the same shape as Fig. 3 of the article: two almost flat and almost identical lines for the control groups, and a much higher and steeper line for the patients, with error bars that are several times longer. Two differences are visible without any calculation. Our points for the patients are higher, especially at the more difficult level, and our line for the patients crosses the vertical axis below zero, whereas the three lines of the article start close together around half a second. Section 1.7 looks for the reason.

## Export the graph as a PDF

The journal asks for a figure of 8 $\times$ 6 inches. The file is written in the folder `figures/` of the repository.

```{r}
#| label: export-pdf
dir.create(here("figures"), showWarnings = FALSE)

ggsave(here("figures", "MT_as_function_of_ID.pdf"),
       plot = graph_MT_ID, width = 8, height = 6, units = "in")
```
